INGRID: Creating Languages in MPS from ANTLR Grammars

### 2016-10-31 chapters/background.tex
Related projects

An MPS language is created by the PE4MPS tool in two steps.
First, the PE parser is used to load an input PE file and build its intermediate representation, which has the form of a tree-like structure consisting of Java objects.
Then, all the concepts (AST nodes) and their aspects are created inside MPS.
The information about the code layout, extracted from the PE file, is used when generating the projectional editor for every AST node.

The main disadvantage of the PE4MPS approach is that it only shifts the tedious manual work from creating projectional editors in the MPS IDE to writing PE grammars in plain text.
Usage of a standard text editor is inferior (and much more error-prone) with respect to MPS, which has been designed for the purpose of specifying the code layout and provides lot of support to developers.
Another limitation of PE4MPS is that it does not implement the full ANTLR syntax.
Every grammar might require non-trivial adjustments before it can be processed using the PE4MPS tool.
To compare, the goals behind our INGRID approach are (1) to enable import of as many languages as possible out-of-the-box with maximal automation and (2) to adopt the full specification of every imported language.

@@

However, the project has also other important limitations.
The tool does not generate any parent-child relationships in the Structure aspect, and it completely neglects the Editor and TextGen aspects.
In particular, it does not support creating of the projectional editor at all.

@@

The main limitation of this approach is the need to perform the last step manually, because MPS currently does not provide any support for automated transformation of a grammar captured by the ANTLRv4 MPS language into the actual target language.

>>>>

The main disadvantage of the PE4MPS approach is that it only shifts the tedious manual work from creating projectional editors in the MPS IDE to writing PE grammars in plain text.
User is forced to input the layout information by extending the grammar manually.
Usage of a standard text editor is inferior (and much more error-prone) with respect to MPS, which has been designed for the purpose of specifying the code layout and provides lot of support to developers.
Another limitation of PE4MPS is that it does not implement the full ANTLR syntax.
Every grammar might require non-trivial adjustments before it can be processed using the PE4MPS tool.
To compare, the goals behind our INGRID approach are (1) to enable import of as many languages as possible out-of-the-box with maximal automation and (2) to adopt the full specification of every imported language.

@@

The major take away from this project is the idea of an intermediate grammar representation inside of MPS which we have also considered.
The advantage of this could be the possibility of extending the ANTLRv4 notation using some graphic tools that MPS has to offer. 
Using the projectional editor, a user could add the layout information in an effective way.
The disadvantage is the dual implementation that is needed.
First to transform the ANTLRv4 text file into the MPS grammar language and secondly to transform this grammar into the target MPS language.

@@
